\section{Background}\label{sec:background}

These notes are not meant to serve as a complete introduction to topos
theory. Therefore, the background section of these notes, rather than
being the first 5 chapters of~\citet{MacLane:92}
or~\citet{Johnstone:14} will contain more or less an accumulation of
definitions and lemmas that we will need. These will be more useful
for ensuring that I have things to reference than for the reader to
learn. It also comes with the moderate advantage that I get to inflict
the my peculiarities upon you dear reader.

Since this is a topos-theoretic proof, it would behoove us to start
with
\begin{defn}\label{defn:background:topos}
  An elementary topos $\Etop{}$ is a category that
  \begin{itemize}
  \item has all finite limits
  \item is cartesian closed
  \item has a subobject classifier
  \end{itemize}
\end{defn}
The topos that we are interested in will be ones that satisfy certain
characteristics making them into a model of ZFC. Modeling the axiom of
choice, the law of the excluded middle, and the existence of infinite
sets in particular present challenges. The next set of definitions are
the categorical analogs of these traits.
\begin{defn}\label{defn:background:nno}
  A natural number object (NNO) is a object $N \in \Ccat$ is an object
  with arrows $s : N \to N$ and $z : 1 \to N$ so that for any object
  $A \in \Ccat$ with $f : 1 \to A$ and $g : A \to A$, there exists a
  unique $h$ so that
  \[
    \begin{tikzcd}
      1 \ar[r, "z"] \ar[dr, swap, "f"] & N \ar[d, dashed, "h"] \ar[r, "s"]& N \ar[d, dashed, "h"]\\
      & A \ar[r, "g"]& A
    \end{tikzcd}
  \]
\end{defn}

\begin{example}
  In the category of sets, $\cat{Set}$, the set of all natural numbers
  $\mathbb{N}$ forms a natural number object.
\end{example}

\begin{example}
  In any presheaf $\presheaf{\Ccat}$, the constant presheaf $A \mapsto
  \mathbb{N}$ forms a natural number object.
\end{example}

\begin{example}
  NNOs are reflected by geometric morphism. In particular, any
  reflective subtopos inherits an NNO from the full topos.
\end{example}

Having categorified the definition of the set of natural numbers, we
now turn to the logical aspects of a topos. The internal logic of a
topos is, in general, intuitionistic and thus validates neither the
law of the excluded middle nor the axiom of choice. We shall need both
of these for the topos we're constructing. Therefore, we turn to
defining what toposes \emph{do} satisfy these principles.

\begin{defn}\label{defn:background:boolean}
  A topos is said to be boolean if the subobject classifier $\Omega$
  forms an internal boolean algebra.
\end{defn}

Booleaness can be captured in several equivalent in useful ways.

\begin{lem}\label{lem:background:boolean}
  The following conditions are equivalent
  \begin{enumerate}
  \item $\Etop$ is boolean
  \item $\neg\neg = 1 : \Omega \to \Omega$
  \item The Heyting algebra $\sub(A)$ for an $A \in \Etop$ is a
    boolean algebra
  \item $\Omega \cong 1 + 1$ with $[\true, \neg \circ \true]$.
  \end{enumerate}
\end{lem}
\begin{proof}\hfill
  \begin{itemize}
  \item $1 \iff 2$\\
    The condition that $\neg\neg a = a$ for all $a$ is equivalent to
    being boolean for any boolean algebra. $a \le \neg \neg a$ in any
    Heyting algebra, therefore all we need to show is that
    $\neg \neg a \le a$. Since $1 \le a \vee \neg a$, it suffices to
    show that $\neg \neg a \wedge (a \vee \neg a) \le a$. Since all
    Heyting algebras are distributive,
    \begin{align*}
      \neg \neg a \wedge (a \vee \neg a) &\le a \Leftrightarrow \\
      (\neg \neg a \wedge a) \vee (\neg \neg a \wedge \neg a) &\le a \Leftrightarrow \\
      a \vee \bot &\le a \Leftrightarrow \\
      a &\le a
    \end{align*}
    For the other direction, it then suffices to show that
    $1 \le \neg \neg (a \vee \neg a)$
    \begin{align*}
      1 &\le \neg \neg (a \vee \neg a) \Leftrightarrow\\
      \neg (a \vee \neg a) &\le 0 \Leftrightarrow\\
      \neg a \wedge \neg \neg a &\le 0 \Leftrightarrow\\
      0 &\le 0 \Leftrightarrow
    \end{align*}
    so we're done.
  \item $1 \iff 3$\\
    We know that the Heyting algebra $\sub(A)$ for any $A \in \Etop$
    corresponds naturally to $\hom(A, \Omega)$. This is a boolean
    algebra precisely when $\Omega$ forms one internally. For the
    reverse direction, if we have $\neg\neg = 1$ in every subobject
    preorder, $\sub(A)$, naturally in $A$, then by Yoneda this holds
    internally to $\Omega$.
  \item $1 \iff 4$\\
    For this, let us first show that $4 \implies 1$. If
    $i = [true, \neg \circ true] : 1 + 1 \cong \Omega$, then it is clear
    that
    \[
      \begin{tikzcd}
        1 + 1 \ar[r, "i"] \ar[d, "tw"] & \Omega \ar[d, "\neg"] \\
        1 + 1 \ar[r, "i"] & \Omega
      \end{tikzcd}
    \]
    commutes where $tw : 1 + 1 \to 1 + 1$ is the canonical twisting
    map. It satisfies the properties that $tw \circ \inl = \inr$ and
    vice versa. Now this means that $i \circ tw \circ i^{-1} = \neg$
    but then
    \[
      \neg\neg = i \circ tw \circ i^{-1} \circ i \circ tw \circ i^{-1} =
      i \circ tw \circ tw \circ i^{-1} = 1
    \]
    so we indeed have that $\Omega$ is an internal boolean
    algebra. Next, we must show that $1 \implies 4$. We know that
    $\true \wedge \neg \true = 0$ so
    $\true \vee \neg \true = \true + \neg \true$ as subobjects of
    $\Omega$. Let us call $m = \true \vee \neg \true$, then
    pictorially, we have that
    \[
      \begin{tikzcd}
        1 + 1 \ar[ddrr, dashed, swap, "m"] & &
        1 \ar[dd, "\true"] \ar[ll]\\
        &&\\
        1 \ar[rr, "\neg \true"] \ar[uu] & & \Omega\\
      \end{tikzcd}
    \]
    However, we know that since $\Omega$ is boolean,
    $a \vee \neg a = 1$. Therefore, $\true \vee \neg \true = 1$. This
    tells us that $1 + 1 \cong \Omega$ as required.
  \end{itemize}
\end{proof}

For our purposes of using a topos to model ZFC, booleaness will be
essential. A boolean topos will validate the law of excluded
middle\footnote{Exercise, show that in a boolean topos
  $\den{\forall x.\ x \vee \neg x}$ holds} which is crucial for
validating the rules of ZFC. There then arises the natural question of
taking an existing topos and modifying it so that it is boolean. This
is easily done using a Lawvere-Tierney topology

\begin{defn}\label{defn:background:lawveretierney}
  A Lawvere-Tierney Topology is a map $j : \Omega \to \Omega$ so that
  \begin{enumerate}
  \item $j \circ \true = \true$
  \item $j \circ j = j$
  \item $\wedge \circ (j, j) = j \circ \wedge$
  \end{enumerate}
\end{defn}
While the subject of Lawvere-Tierney topologies on an elementary topos
is a rich one, we shall limit ourselves to the essentials for our
purposes here. It will be important to know that given a
Lawvere-Tierney topology we can internally define a notion of
``sheaf'' which matches up to that of a sheaf on a Grothendieck
topology. In order to do this, first we define
\begin{defn}\label{defn:background:closure}
  For a topology $j$ on $\Etop$, the closure of $A \mono B$ is the
  subobject of $B$ classified by $j \circ \charmap(A)$, $\bar{A}$. A
  subobject is said to be closed if $\bar{A} = A$.
\end{defn}

We can define an alternative subobject classifier which internalizes
a Lawvere-Tierney topology as
\[
  \begin{tikzcd}
    \Omega_j \ar[r, dashed] &
    \Omega \ar[r, shift left=0.5em, "1"]
    \ar[r, shift right=0.3em, swap, "j"] &
    \Omega
  \end{tikzcd}
\]
This has the nice property that it classifies closed subobjects.
\begin{lem}\label{lem:background:closedfactors}
  If $A \mono B$ is a closed subobject under $j$, then $\charmap(A)$
  factors through $\Omega_j$.
\end{lem}
\begin{proof}
  Suppose that $A$ is a closed subobject, then we know that
  $\charmap(A) \circ j = \charmap(A)$ by definition. Therefore,
  \[
    \begin{tikzcd}
      A \ar[d, dashed] \ar[dr, "\charmap(A)"] & &\\
      \Omega_j \ar[r]&
      \Omega \ar[r, shift left=0.3em, "1"] \ar[r, shift right = 0.3em, swap, "j"] & \Omega
    \end{tikzcd}
  \]
  by the universal property of an equalizer.
\end{proof}

\begin{defn}\label{defn:background:dense}
  A subobject $A \mono B$ is said to be dense if $\bar{A} \cong B$
\end{defn}

We are now in a position to define the internalization of sheaves. The
inspiration for this definition comes from the fact that in a
Grothendieck topology, a sieve is a dense subfunctor of
$\yoneda(C)$. We therefore can define sheaves purely in terms of dense
subobjects.

\begin{defn}\label{defn:background:sheaves}
  An object $F \in \Etop$ is said to be a sheaf if for every dense
  subobject $A \mono B$, a morphism $f : A \to F$ has a unique
  extension $g : B \to F$.
  \[
    \begin{tikzcd}
      A \ar[d, rightarrowtail, swap, "m"] \ar[r, "f"] & F\\
      B \ar[ur, dashed, swap, "g"]
    \end{tikzcd}
  \]
\end{defn}

Sheaves enjoy several nice properties with regards to closure which
will be useful for establishing some facts later on.
\begin{lem}\label{lem:background:sheavesclosed}
  For any $A \mono B$ if $A$ is a sheaf then $A$ is closed.
\end{lem}
\begin{proof}
  We know that by definition that the mono $m : A \mono \bar{A}$ is
  dense and since $A$ is a sheaf we must have
  \[
    \begin{tikzcd}
      A \ar[d, rightarrowtail, "m"] \ar[r, equals] & A\\
      B \ar[ur, dashed]
    \end{tikzcd}
  \]
  so in particular $m$ must be an isomorphism.
\end{proof}

Finally, while the proof is too far out of scope for these notes, it
is important to note that sheaves on a topology $j$ form a full
subcategory of a topos, $\Etop$. This category, denoted
$\sheaves[j]{\Etop}$, is in fact a reflective subcategory with
\[
  \begin{tikzcd}
    \sheaves[j]{\Etop} \ar[rr, bend left, "\iota"] & \top &
    \makebox[0pt]{\phantom{\textbf{Sh}}$\Etop$}\phantom{\sheaves[j]{\Etop}}
    \ar[ll, bend left, "\sheafify"]
  \end{tikzcd}
\]
The curious reader is referred to either~\citet{MacLane:92}
or~\citet{Johnstone:14}. Additionally Jon Sterling gave a talk on this
and ought to distribute his notes. Please go bother him if you really
want a proof. Now in our case, we shall be interested in one
particular topology, the double negation topology.
\begin{example}\label{ex:background:doublenegation}
  For any topos, $\neg\neg : \Omega \to \Omega$ forms a topology.
\end{example}
The useful property of the double-negation topology for our purposes
is that $\sheaves[\neg\neg]{\Etop}$ will always form a boolean
subtopos of $\Etop$. Later on, we will use this to construct a boolean
topos out of a topos modeling the forcing construction we wish to
implement.

\begin{lem}\label{lem:background:notnotboolean}
  $\sheaves[\neg\neg]{\Etop}$ is boolean.
\end{lem}
\begin{proof}
  By~\ref{lem:background:boolean} it suffices to show that for each
  sheaf $A$ that $\sub(A)$ is a boolean lattice. If $B \mono A$ is a
  subobject of $A$, it must be that $B$ closed as it too must be a
  sheaf by~\ref{lem:background:sheavesclosed}. But closure in under
  $\neg\neg$ means precisely that $\neg \neg B = B$ so $\sub(A)$ is
  boolean as required.
\end{proof}

An interesting aside at this point is that the double-negation
topology on a presheaf topos has a well known Grothendieck analog: the
dense topology. That is, the topology given by
\[
  J(C) = \{ S \mid \forall f : A \to C.\ \exists h.\ f \circ h \in S\}
\]
I will not prove this but will use the phrase ``dense topology'' and
``double negation topology'' interchangeably in the
notes\footnote{It's because I'm a mean-spirited person}.

We next turn to the topos-theoretic analog of the axiom of
choice. Here we are presented with two possible ways. The first is a
direct formulation of the principle from the category of sets. We wish
to generalize \emph{every surjection has a section}. We can generalize
this to a topos by replacing surjection with epimorphism to get

\begin{defn}\label{defn:background:aoc}
  A topos satisfies the axiom of choice if every $e : A \epi B$ has a
  section $s : B \to A$ so that $es = 1$
\end{defn}

However, it is often preferable to state a version termed the internal
axiom of choice but the differences are out of scope for these
notes. I will settle for merely stating it.

\begin{defn}\label{defn:background:iaoc}
  A topos satisfies the internal axiom of choice if $(-)^A$ preserves
  epimorphisms.
\end{defn}

\begin{remark}
  Any topos that satisfies the axiom of choice satisfies the internal
  version. Any topos that satisfies the internal version is
  boolean. This is due to~\citet{Diaconescu:75}.
\end{remark}

This characterization of the axiom of choice is awkward to work with
however in many toposes. A cleaner characterization can be given in
terms of the what objects \emph{generate} the topos.

\begin{defn}\label{defn:background:generate}
  A collection of objects $S$ is said to generate a category $\Ccat$
  if for any parallel arrows $f \neq g : A \to B$, there exists an
  arrow $h : C \to A$ for some $C \in S$ so that
  $fh \neq gh : C \to B$. In particular, if a topos is generated by
  $1$ and is nontrivial then it is \emph{well-pointed}.
\end{defn}

Generation by a collection of objects captures a large number of
logical principles that we will need for our topos. Most important
among these is that for a boolean Grothendieck topos generated by
subobjects of 1, we may automatically derive the validity of the axiom
of choice. This is because in a Grothendieck topos the subobject
preorders are all \emph{complete} partial orders. We demonstrate this
in the next lemma.

\begin{lem}\label{lem:background:completealgebratoaoc}
  If a boolean topos $\Etop{}$, is generated by subobjects of $1$ and
  has complete boolean algebras for subobject posets then $\Etop{}$
  satisfies the axiom of choice.
\end{lem}
\begin{proof}
  We wish to show that $e : A \epi B$ has a section $B \to A$. We then
  wish to use Zorn's lemma to find a maximal subobject $m : M \mono B$
  so that $e$ has a section $s : M \to A$. That is, so that $es = m$.
  In order to apply Zorn's lemma we wish to show that the subset of
  $\sub(B)$ is closed under chains. First we note that any such
  section is necessarily mono. This is because if $sf = sg$, then we
  must have $esf = esg$ so $mf = mg$ so $f = g$ as $m$ is mono. Now
  let us show the chain condition.

  Suppose that $(M_i)_i \subseteq \sub(B)$ so that for each $M_i$, we
  have a section $s_i$ and this forms a chain. Now we note that each
  $M_i$ is a subobject of $B \times A$ under
  $\langle m_i, s_i \rangle$. Since we assumed that $\Etop{}$ had
  complete boolean algebras for subobject lattices, let us take
  $M = \bigvee_i M_i$ taken as subobjects of $B \times A$. Now this
  tells us that $M \mono B \times A$ so we must have $m : M \to B$ and
  $s : M \to A$. We must show that $e \circ s = m$ now.

  Let us note that $A$ is a subobject of $B \times A$ under
  $\langle e, id \rangle$. Furthermore, by $s$ we have that
  $M_i \le A$ for all $i$. Therefore, by the definition of a least
  upper bound, we must have $M \le A$. Therefore, we have a mono so
  that
  \[
    \begin{tikzcd}
      M \ar[rr, rightarrowtail, "s'"] \ar[dr, swap, rightarrowtail, "{\langle m, s \rangle}"] &&
      A \ar[dl, rightarrowtail, "{\langle e, id \rangle}"]\\
      &B \times A&
    \end{tikzcd}
  \]
  this tells us that $s' = s$ so we have our desired section. Now we
  want to conclude that $m : M \to B$ is indeed a mono, however, all
  we know is that $\langle m, s \rangle$ is a mono. Therefore, we want
  to show that
  \[
    (x, x', y \mid m(x) = y \wedge m(x') = y \vdash x = x')
  \]
  in the internal logic. Since $m' = \langle m, s \rangle$ is a mono,
  we know that
  \[
    (x, x', z : B \times A \mid m'(x) = z \wedge m(x') = z \vdash x = x')
  \]
  Therefore, if we can show that
  \[
    (a, a', b \times M(b, a), M(b, a') \vdash a = a')
  \]
  then we could rewrite the assumption that $m'$ is mono to
  \[
    (x, x', b : B \mid \pi_1(m'(x)) = b \wedge \pi_1(m(x')) = b \vdash x = x')
  \]
  and since $m = \pi_1 \circ m'$ by definition, this gives us our
  desired result. We know that since each $m_i$ is mono, that if
  \[
    (x, x', b \mid m_i(x) = b \wedge m_i(x') = b \vdash x = x')
  \]
  Furthermore, this tells us that if
  \[
    (x, x', b, a, a' \mid (\langle m_i, s_1 \rangle)(x) = (b, a) \wedge
    (\langle m_i, s_1 \rangle)(x') = (b, a') \vdash x = x')
  \]
  so by functionality, since we have
  $f(x) = (b, a) \wedge f(x) = (b, a')$ we must have $a = a'$ as
  well. We know that for all $i$.
  \[
    (a, a', b \times M_i(b, a), M_i(b, a') \vdash a = a')
  \]
  Now, stepping outside the internal logic, we wish to show that
  \[
    \begin{tikzcd}
      {\pi_{12}^*(M) \wedge \pi_{13}^*(M)} \ar[dr] \ar[rr, rightarrowtail] & &
      {B \times A} \ar[ld, "{1 \times \delta}"]\\
      &{B \times A \times A}&
    \end{tikzcd}
  \]
  Let us now work with $\sub(B \times A \times A)$.
  \begin{align*}
    \pi_{12}^*(M) \wedge \pi_{13}^*(M) &= \pi_{12}^*(\bigvee_i M_i) \wedge \pi_{13}^*(\bigvee_i M_i)\\
      &= \bigvee_i \pi_{12}^*(M_i) \wedge \bigvee_i \pi_{13}^*(M_i)\\
      &= \bigvee_i \bigvee_j \pi_{12}^*(M_i) \wedge \pi_{13}^*(M_j)\\
      &= \bigvee_i \pi_{12}^*(M_i) \wedge \pi_{13}^*(M_i)\\
      &= \bigvee_i \delta\\
      &= \delta
  \end{align*}
  Where we have made use of the fact that clearly
  \[
    \bigvee_i \pi_{12}^*(M_i) \wedge \pi_{13}^*(M_i) \le
    \bigvee_i \bigvee_j \pi_{12}^*(M_i) \wedge \pi_{13}^*(M_j)
  \]
  but since $M_i$ forms a chain, supposing that $i \le j$ without loss
  of generality, we then have that
  \[
    \pi_{12}^*(M_i) \wedge \pi_{13}^*(M_j) \le \pi_{12}^*(M_j) \wedge \pi_{13}^*(M_j)
  \]
  This means that the reverse inclusion holds. All told, this means
  that we have indeed constructed an $M$ dominating the chain with a
  partial section for $e$ as required. Thus, Zorn's lemma gives us a
  maximal such subobject.

  Now suppose we have such a maximal subobject $m : M \mono B$ with a
  section $s : M \to A$. Let suppose that $M \neq A$, for it is then
  we're done. We note that $M$ must have a nonzero complement as it is
  not $B$. Since $\neg M$ is nonzero, it has two different subobjects
  of its own, $0$ and $1$. Thus $\neg M$ has two different
  characteristic maps and since subobjects of 1 generate, for some
  $U \mono 1$ we must have $t : U \to \neg M$. Consider
  \[
    \begin{tikzcd}
      A' \ar[d, "e'", twoheadrightarrow] \ar[rr] & & A \ar[d, twoheadrightarrow, "e"]\\
      V \ar[r, rightarrowtail, "t"] & \neg M \ar[r, rightarrowtail] & B
    \end{tikzcd}
  \]
  Next, we note that $V \neq 0$ because it has an arrow to a nonzero
  object. From this, it follows that $A'$ must not be 0 as $A' \epi V$
  and
  $\begin{tikzcd}
    V \ar[r, "f", shift left=0.2em] \ar[r, "g", swap, shift right=0.2em] & \Omega
  \end{tikzcd}$
  for some $f \neq g$. If $A'$ was zero then it must be that $fe = ge$
  which would give us $f = g$, a contradiction.

  Now since $A'$ is nonzero, we can once again get a subterminal
  object which maps into it, $W \mono 1$ with $w : W \to A'$. Now we
  immediately have $e'w : W \to V$ so it follows that $W \mono V$. Now
  we take the image factorization of $\Im(te'w) = e'' : W \to t(W)$.
  Now since $W \mono 1$, it must be that $e''$ is also mono and since
  $\Etop$ is balanced, we then have that $e''$ is an iso. Now as
  $t(W) \mono \neg M$, it must be that $t(W) \wedge M = 0$ so
  $t(W) \vee M = t(W) + M$. Moreover, $we''^{-1} : t(W) \to A'$ so we
  know that $t(W) \to A$ so we can form a section $M + t(W) \to A$.
  This contradicts the maximality of $M$.
\end{proof}

With this we are in a position to start our proof because a topos
which satisfies the AoC and has a NNO is powerful enough to provide a
model of ZFC that we will use to validate the independence of the
continuum hypothesis.
